---
title: "Advanced Marketing Modelling -- Jobs"
author: "Arshia Gupta, Elias Ganal, Yaonan Lyu, Ziheng Qiu"
date: "2025-01-22"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Load packages
```{r}
library(dplyr)
library(ggplot2)
library(corrplot)
library(psych)
library(data.table)
library(dendextend)
library(factoextra)
library(ggsignif)
library(plotrix) # For Std.Errors
```



```{r}
# Read the csv file
setwd("/Users/arshia/Group Jobs")
microvan <- read.csv("microvan.csv", sep = ";")
```

## Data Preparation
```{r}
# Check duplicated values
duplicates <- microvan[duplicated(microvan), ]
sum(duplicated(microvan))
#Check missing values 
sum(is.na(microvan))
```

### Descriptive analysis
```{r}
psych::describe(microvan[-1], fast = TRUE)
```

### Distribution of dependant variable - Concept Liking
```{r}
ggplot(microvan, aes(x = mvliking)) +
  geom_histogram(binwidth = 1, fill = "skyblue", color = "black") +
  labs(title = "Distribution of Concept Liking", x = "mvliking", y = "Frequency") +
  theme_minimal()
```

### Standardize variables
```{r}
vars <- scale(microvan[,-1])

psych::describe(vars, fast = T)
```

### Correlation Matrix of all variables
```{r}
corrplot( cor(vars), 
         method = "number", 
         type = "upper", 
         order = "hclust", # reorder by the size of the correlation coefficients
         tl.cex = 0.5, # font size of the variable labels
         tl.col = "black", # color of the variable labels
         tl.srt = 45, # rotation angle for the variable labels
         number.cex = 0.25 # font size of the coefficients
)
```

## Factor Analysis
### Select variables(eliminate all demographic variables)
```{r}
FAvars <- microvan[,c("kidtrans",
                      "miniboxy",
                      "lthrbetr",	
                      "secbiggr",
                      "safeimpt",	
                      "buyhghnd",
                      "pricqual",
                      "prmsound",
                      "perfimpt",
                      "tkvacatn",
                      "noparkrm",
                      "homlrgst",	
                      "envrminr",
                      "needbetw",
                      "suvcmpct",
                      "next2str",
                      "carefmny",
                      "shdcarpl",
                      "imprtapp",
                      "lk4whldr",
                      "kidsbulk",
                      "wntguzlr",
                      "nordtrps",
                      "stylclth",
                      "strngwrn",
                      "passnimp",
                      "twoincom",
                      "nohummer",
                      "aftrschl",
                      "accesfun")]

cor = cor(FAvars)

library(corrplot)
corrplot(cor, 
         method = "number", 
         type = "upper", 
         order = "hclust", # reorder by the size of the correlation coefficients
         tl.cex = 0.5, # font size of the variable labels
         tl.col = "black", # color of the variable labels
         tl.srt = 45, # rotation angle for the variable labels
         number.cex = 0.35 # font size of the coefficients
)

```

```{r}
# KMO and MSA(i)
psych::KMO(cor) 
```


```{r}
# Bartlett’s test for Sphericity
psych::cortest.bartlett(cor,
                 n = dim(microvan)[1]) # number of rows
```

### Find the eigenvalues
```{r}
EV = eigen(cor)$values
EV
```

```{r}
# The individual percentage of variance:
EV/length(EV)
```

```{r}
#scree plot
scree(cor, pc = TRUE, factors = FALSE)
```

```{r}
# Shares for the cumulative variance explained
plot(cumsum(EV/length(EV)), 
     type = "o", # type of plot: "o" for points and lines 'overplotted'
     col = "darkblue",
     pch = 16, # plot symbol: 16 = filled circle
     cex = 1, # size of plot symbols
     xlab = "Number of factors", # a title for the x axis
     ylab = "Cumulative variance explained", # a title for the y axis
     lwd = 2) # line width
abline(v = 3, lwd = 1.5, col = "grey") # draw a vertical line at v = ...
abline(h = 0.6, lwd = 1.5, col = "grey", lty = 2) # draw a horizontal line at h = ...

```

### Execute the PCA
```{r}
PCA <- principal(r = cor, 
                 nfactors = 4, 
                 rotate="varimax",
                 scores = TRUE)
print(PCA, 
      digits = 3, # to round numbers to the third digit 
      cut = 0.35, # to show only values > 0.35
      sort = TRUE # to sort rows by loading size
)
```

```{r}
# Visualize the PCA
fa.diagram(PCA, cex = 0.2, cut = 0.6)
```

```{r}
# The matrix of loadings:
library(data.table)
L <- as.data.table(unclass(PCA$loadings), keep.rownames = T)

plot(x = L$RC1, y = L$RC2, 
     col ="darkblue",  
     pch = 16,        # plot symbol: 16 = filled circle
     cex = 1,         # size of plot symbols
     xlab = "Factor 1",    # a title for the x axis
     ylab = "Factor 2",    # a title for the y axis
     xlim = c(-1.5,1.5),  # x axis values from -1 to 1
     ylim = c(-1.5,1.5))  # y axis values from -1 to 1

# add point labels
text(L$RC1, L$RC2, 
     labels = L$rn,
     pos = 3, 
     cex = 0.8, 
     col = "darkred")

# add vertical and horizontal lines
abline(h = 0, lwd = 2, col = "grey") # draw a horizontal line at h = 0
abline(v = 0, lwd = 2, col = "grey") # draw a vertical line at v = 0
```

```{r}
PCA$r.scores
```

### Perceptual map
```{r}
# extract  scores of principal components
PCA.scores = factor.scores(FAvars, unclass(PCA$loadings))$scores
```

```{r}
#standardized value
psych::describe(PCA.scores, fast = TRUE)
```

```{r}
## Add factor scores to the initial dataset
microvan.scores <- cbind(microvan, PCA.scores)

plot(x = microvan.scores$RC1, 
     y = microvan.scores$RC2,
     xlab = "Factor 1", ylab = "Factor 2",
     xlim = c(-3, 3), ylim = c(-3, 3), 
     pch = 16, cex = 1, col = "blue")

abline(h = 0, col = "grey")
abline(v = 0, col = "grey")

# add point labels
text(x = microvan.scores$RC1, 
     y = microvan.scores$RC2, 
     labels = microvan.scores$state,
     cex = 1, 
     adj = 1.2, 
     col = "black")
```

## Cluster Analysis

```{r}
data.dist<-dist(microvan.scores[,c("RC1", "RC2", "RC3", "RC4")],
                method = "euclidean") # one set of variables selected
as.matrix(data.dist)[1:5,1:5] # Visualize distances among the first 5 observations
```


### Number of clusters
```{r}
cluster.ward <- hclust(data.dist, method ="ward.D2")


plot(as.dendrogram(cluster.ward), # select the cluster solution
     ylab = "Distance",
     main = "Dendrogram",          # specify the plot title
     cex = 0.3)                    # specify the label size
```

```{r}
n_clusters = 4
plot(set(as.dendrogram(cluster.ward), # select the cluster solution
         "branches_k_color",       # color the cluster branches for the number of clusters (k = 4)
         k = n_clusters),                   # specify the number of clusters
     ylab = "Distance",
     main = "Dendrogram",          # specify the plot title
     cex = 0.3)                    # specify the label size
rect.hclust(cluster.ward, k = n_clusters, border = "darkred")  # draw red borders around the clusters

```

```{r}
factoextra::fviz_nbclust(x = microvan.scores[,c("RC1", "RC2", "RC3", "RC4")], # data
                         FUN = hcut, # hierarchical cluster
                         method = "wss") 
```

```{r}
cluster.ward.segment <- cutree(cluster.ward, k=n_clusters)
```

```{r}
table(cluster.ward.segment) # cluster sizes
```


```{r}
table(cluster.ward.segment)/nrow(microvan.scores) # proportions of the sample
```

### Description of clusters
```{r}
microvan$cluster <- cutree(cluster.ward, k=n_clusters) # to store the cluster solution in the initial data
```

```{r}
# Average values of variables for each cluster
aggregate(microvan[-1],
          by = list(cluster = microvan$cluster), 
          FUN = mean)
```

```{r}
# Generate the summarized dataset first
summarized_data <- microvan[-1] %>% 
  group_by(cluster) %>% 
  summarise_all(.funs = list(~mean(., na.rm = T))) %>% 
  tidyr::gather(Variable, Mean, mvliking:recycle) %>% 
  left_join(
    microvan[-1] %>% 
      group_by(cluster) %>% 
      summarise_all(.funs = list(~std.error(., na.rm = T))) %>% 
      tidyr::gather(Variable, SE, mvliking:recycle),
    by = c("cluster", "Variable")
  )

# Define variable groups
variable_groups <- list(
  group1 = c("kidtrans", "miniboxy", "lthrbetr", "secbiggr", "safeimpt", "buyhghnd", "pricqual", "prmsound"),
  group2 = c("perfimpt", "tkvacatn", "noparkrm", "homlrgst", "envrminr", "needbetw", "suvcmpct", "next2str"),
  group3 = c("carefmny", "shdcarpl", "imprtapp", "lk4whldr", "kidsbulk", "wntguzlr", "nordtrps", "stylclth"),
  group4 = c("strngwrn", "passnimp", "twoincom", "nohummer", "aftrschl", "accesfun", "mvliking"),
  group5 = c("age", "income", "miles", "numkids", "female", "educ", "recycle")
)

# Function to filter data and create a plot
create_plot <- function(data, variable_subset, title_suffix) {
  data %>%
    filter(Variable %in% variable_subset) %>%
    ggplot(aes(y = Mean, x = cluster)) +
    facet_wrap(~ Variable, scales = "free", ncol = 4) +
    geom_bar(stat = "identity", fill = "#bac5db") +
    geom_errorbar(aes(ymin = Mean - 1.96 * SE, ymax = Mean + 1.96 * SE),
                  width = 0.2, position = position_dodge(0.9)) +
    labs(x = "Cluster", y = "Mean + 95%-Confidence Interval") +
    theme_minimal(base_size = 17) +
    theme(axis.text.x = element_text(angle = 0)) +
    ggtitle(label = paste("Descriptive Statistics for Clusters -", title_suffix))
}

# Create plots for each group
plot1 <- create_plot(summarized_data, variable_groups$group1, "Group 1")
plot2 <- create_plot(summarized_data, variable_groups$group2, "Group 2")
plot3 <- create_plot(summarized_data, variable_groups$group3, "Group 3")
plot4 <- create_plot(summarized_data, variable_groups$group4, "Group 4")
plot5 <- create_plot(summarized_data, variable_groups$group5, "Group 5")

# View the plots
print(plot1)
print(plot2)
print(plot3)
print(plot4)
print(plot5)

```

## Regression
### logistics regression
```{r}
# Determine the 75th percentile
threshold <- quantile(microvan.scores$mvliking, 0.75, na.rm = TRUE)

# Create the binary variable
microvan.scores$mvliking_binary <- ifelse(microvan.scores$mvliking > threshold, 1, 0)

# Check the distribution of the binary variable
table(microvan.scores$mvliking_binary)
```

### Null model
```{r}
m0 <- glm(mvliking_binary ~1, data = microvan.scores, family = "binomial")
summary(m0)
```

### Full model
```{r}
# Fit the logistic regression model
  
m1 <- glm(mvliking_binary ~ RC1 + RC2 + RC3 + RC4+age+income+miles+numkids+female+educ+recycle, 
                      data = microvan.scores, family = "binomial")

summary(m1)
```

### Final model
```{r}
  
m2 <- glm(mvliking_binary ~  RC2 + RC4+age+miles+female+educ, 
                      data = microvan.scores, family = "binomial")

summary(m2)
```

```{r}
round(exp(coef(m2)), 3)
```
```



